Effects of bariatric surgery and weight loss on resting state functional connectivity

### 2021-01-27 manuscript/draft_rsfmri.Rmd
modification of structure & figures after re-categorization of one participant

```{r "knit setup", include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE) # fig.height=2, fig.width=4)
options(knitr.table.format = "latex") 
options(knitr.kable.NA = '') # hide NA in tables
# set working directory of knit (RMarkdown) to directory of Rproj
opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
setwd(knitr::opts_knit$get("root.dir"))
```

@@

Resting state functional magnetic resonance imaging (rsfMRI) is a powerful technique to investigate the dynamic organization of the brain via the correlation of spontaneous neural activity. Anatomically distinct brain regions show functional connectivity (FC), i.e. correlated neural activity over time, in the absence of a cognitive task. Crucially, these brain networks correspond to networks activated during specific cognitive processes, e.g. visual, auditory and somatosensory processing, or higher-order cognitive functions, i.e. attention and executive control [@Smith_2011], and can thus be seen as default configuration of the brain. In the context of bariatric surgery, the default mode network, a higher-order network, involved in interoception and governing shifts between external-internal processes, and the reward network, processing hedonic value and internal motivation, are promising candidates to mediate altered gut-brain signaling.

@@

Higher BMI and obesity have been associated with differences in DMN FC in several resting state and task-based fMRI studies [@Beyer_2017; @Kullmann_2012; @Chao_2018; @Wijngaarden_2015, @Borowitz_2020; @Ding_2020; @Doucet_2017; @Sadler_2018]. 

@@

```{r "Load data", include=FALSE}
source("report/create_sample_df.R")
final = create_sample_df(group = "all", tp = "all")

@@

```{r figBMIdescr, fig.align='center', out.width="90%", fig.cap="BMI over time.", echo=FALSE}
# eval = TRUE, fig.asp = 0.7, fig.width = 6
source("report/figures.R")

fig_BMIdescr <- mk_figBMIdescr(final)
fig_BMIdescr

@@

<!-- MODEL SETUP --> In a first group of models, we tested the interaction of time and group $FC = group + time + group*time$ as specified in the preregistration (*We will use the SwE toolbox (implemented in SPM 12.7486 run in MATLAB version > 9.0) to calculate the group-by-timepoint interaction*)
We did not perform the analysis as preregistered (i.e. without covariates and additional sensitivity analyses), but adjusted for baseline age and sex (Model 1a), and baseline age, sex and the logarithm of mean FD (Model 1b). 
<!-- DESIGN MATRIX, vielleicht beide auch als Abbildung kurz zeigen? --> In this first model, time was represented as factor because we did not consider the effect of time to be strictly linear, and thus continuous modeling of time, inappropriate (see Figure \@ref(fig:figDesignmatrix)). The resulting factorial design <!-- check: flexible factorial? --> contained one regressor for each time point per group.
```{r figDesignmatrix, fig.align = 'center', fig.ncol = 2, out.width = "50%", fig.subcap=c('Design matrix with time as categorial factor.','Design matrix with time as continuous variable.'), fig.cap="Possible design matrices of time-by-group model.", echo=FALSE}
# fig.show="hold",  out.width="40%", fig.height = 8,
source("report/figures.R")
PlotList <- mk_figDesignMatrix()

@@

## Quality control
We performed different quality controls for the rsfMRI data.
First, we inspected parameters of head motion (mFD) and a measure of whole-brain signal variation which is reflective of noise in the data (DVARS).
As described in @Beyer_2020, there was a group-by-time interaction on mFD. The intervention group reduced in head motion in the follow-up assessments compared to the control group. 
Further, DVARS did not qualitatively differ between groups and time points, and the correlation of DVARS and mFD remained similar over time points.
We concluded that even though there was a decrease in head motion in the intervention versus control group, the impact of head motion on imaging data remained similar over the study period in both groups.
```{r tableQcfc, echo=FALSE}
res <- read.csv("./report/FDFCcorrelations.csv")
rownames(res) <- c("minimally processed", "AROMA", "AROMA + CC", "AROMA + CC + GSR")
colnames(res) <- c('mean FD-QC', 'median FD-QC', 'sig. vertex', 'sig. vertex BH', 'distance-FD-QC', 'pval')

@@

```{r FDFC correlation, eval=FALSE, echo=FALSE}
source("/data/pt_02161/Analysis/Preprocessing/qa/rs_qa/group_level_QA/QC_FC_correlations/calc_qc_fc_correlations.R")
FDFC=calc_qc_fc(final)
write.csv(FDFC,"./report/FDFCcorrelations.csv", row.names = FALSE)
```

```{r figNetworks, echo=FALSE, out.width="49%", out.height="30%", fig.cap="Resting state networks based on AROMA+CC+GSR input data, one-sample t-tests adjusting for age and sex over all timepoints and participants with bootstrapped clusterwise inference (p FWE<0.05)", fig.subcap=c("DMN seeded from precuneus", "Reward network seeded from Nucleus Accumbens", "tSNR differences between regions"),fig.show='hold',fig.align='center'}
#networks=c("./report/DMN_thresholded_for_clusterFWE005.png","./report/Rew_thresholded_for_clusterFWE005.png")
networks=c("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/DMN_thresholded_for_clusterFWE005.png", "/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/Rew_thresholded_for_clusterFWE005.png", "/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/tsnr.jpeg")
knitr::include_graphics(networks)
```

Figure \@ref(fig:figNetworks) shows the DMN and reward network average maps which we used for extracting aggregated FC values. Both networks included the typically described brain regions (i.e. medial prefrontal cortex, parietal lobules and posterior hippocampus for the DMN, and anterior cingulate, right amygdala and ventromedial-prefrontal cortex for the reward network). 
Yet, the reward network was less pronounced (overall lower T-values) and symmetric (no significant connectivity with left amygdala), which might be due to lower SNR in the seeded brain region (see Figure \@ref(fig:figNetworks).
```{r prepare tsnr ROI plot, eval=FALSE}
final$subj.ID_tp=paste0(final$subj.ID,'_', final$tp)

@@

p<-ggplot(aes(as.factor(roi), mean),data=final_tsnr) +
  geom_violin(aes(fill=roi)) + geom_jitter(height = 0, width = 0.1) + xlab("Region of interest") + ylab("average tSNR") + 
  scale_fill_manual(values=wes_palette("Darjeeling1",5,type="discrete")[3:4]) +
  theme(axis.text=element_text(size=10),
  axis.title=element_text(size=12),
  strip.text = element_text(size=12),
  legend.position="")
ggsave("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/tsnr.jpeg", p)
```

@@

<!-- Brain mask: Group*Time Interaction [CORRECTED] -->
There was no interaction of group and time point regarding the resting state connectivity for neither the PCC nor the NAcc seeds against our initial hypothesis. There also was no significant main effect for any of the covariates of interest (time, group) in clusterwise inference with FWE correction.

####  Model 1b (with adjustment for age and sex, meanFD)
Similarly, there were no significant clusters when including log mean FD. Results did not differ between AROMA+CC and AROMA+CC+GSR preprocessing pipelines.

@@

## Interaction effect group by time on aggregated FC (preregistered analysis)
There was no significant time by group interaction for aggregated DMN and reward network FC (see Figures \@ref(fig:figaggDMN) and \@ref(fig:figaggRew) and supplements for a detailed summary of the models) 
```{r load agg FC data and prepare, echo=FALSE}
agg_FC_DMN=read.table("/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/aggFC/DMN/agg_FC_DMN.txt")
agg_FC_DMN_ID=read.table("/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/aggFC/DMN/agg_FC_DMN_ID.txt")
agg_FC_Rew_ID=read.table("/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/aggFC/Rew/agg_FC_Rew_ID.txt")
agg_FC_Rew=read.table("/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/aggFC/Rew/agg_FC_Rew.txt")
for (i in c(1:nrow(agg_FC_DMN_ID))){
  agg_FC_Rew[i,"subj.ID_tp"]=strsplit(base::strsplit(as.character(agg_FC_Rew_ID[i,"V1"]),'/')[[1]][10],'_nacc')[[1]][1]
  agg_FC_DMN[i,"subj.ID_tp"]=strsplit(base::strsplit(as.character(agg_FC_DMN_ID[i,"V1"]),'/')[[1]][10],'_pcc')[[1]][1]
}

@@

#calculate mean FD over timepoints
final_FC.meanFD=tapply(X=final_FC$logmFD,
                        INDEX=final_FC$subj.ID, FUN=mean, na.rm=TRUE)
final_FC$mean.logmFD=
  final_FC.meanFD[as.numeric(final_FC$subj.ID)]
final_FC$within.logmFD=
  final_FC$logmFD-final_FC$mean.logmFD
```

@@

```{r figaggDMN, echo=FALSE}
fig_DMNconn<-mk_figDMNdescr(final_FC)
fig_DMNconn
```

@@

R1 = lme4::lmer(mean_Rew_conn ~ tp*group_factor + Age_BL + Sex + logmFD +(1|subj.ID), data=final_FC)
R0 = lme4::lmer(mean_Rew_conn ~ tp + group_factor + Age_BL + Sex +logmFD +(1|subj.ID), data=final_FC)
anvRew3tp=anova(R1,R0)
```
```{r Rew3tp results, caption="DMN connectivity compared bl to fu", results="asis", eval=FALSE}
#tab_model(R1,R0)
texreg(list(R0,R1), reorder.coef=c(1, 2, 3, 4, 8,9,5,6,7),custom.coef.names=c("Intercept (BL, NBARS)", "Timepoint FU", "Timepoint FU2", "Group BARS", "Age", "Sex","logmFD","Timepoint FU x Group", "Timepoint FU2 x Group"), custom.model.names = c("R0", "R1"), float.pos = "h",  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
```
```{r figaggRew, echo=FALSE}
fig_Rewconn<-mk_figRewdescr(final_FC)
fig_Rewconn
```

## Exploratory Analyses
### Effect of mean and change in BMI 
#### Whole brain analysis
Next, we tested for significant effects of the average BMI and the individual change in BMI longitudinally across the three time points.
<!-- ### DMN -->
<!-- cc -->
Using the images preprocessed with AROMA+CC, we found a significant negative association of average BMI and DMN connectivity in the lingual gyrus, mid orbital gyrus and temporal gyrus, which remained stable after including mFD as nuisance covariate. When using GSR preprocessed data, none of the clusters was significantly associated with average BMI.
NAcc connectivity was not significantly related to neither mean nor change in BMI, for either of the preprocessing schemes. 
Results with significant cluster co-activation are depicted in Table \@ref(tab:tableBMImodel).
To disentangle effects of change in BMI and logmFD in this dataset, we also split mFD into average mFD and change in mFD. Age and sex remained as nuisance covariates.
<!-- ### DMN -->
<!-- cc --> Average BMI remained significantly negatively associated with the three clusters above when splitting FD into its components. Results are detailed in Table \@ref(tab:tableBMIFDmodel).
Test statistics of these clusters differed only slightly from the effect found for average BMI when using mFD as nuisance covariate.
<!-- gsr -->As before, the effect did not survive when using GSR-corrected data.

<!-- ### Reward -->
<!-- cc -->
For the reward network, we found a significant negative association of change in BMI with FC between NAcc and a cluster in the posterior-medial frontal region, i.e. supplementary motor area (see Table \@ref(tab:tableBMIFDmodel)). Voxel activation at local maximum within this cluster was significant at peak level after FWE-correction. 
Moreover, FC between the NAcc and one cluster in a more posterior region, i.e. motor cortex, was significantly positively related to average mFD.
When running the model with GSR-preprocessed data, the associations between change in BMI and FC did not survive statistical thresholds. The relationship with average mFD was confirmed by analyses after GSR but cluster size shrank.

@@

p_mean_BMI=anvmeanBMI$`Pr(>Chisq)`
```
The association also remained significant when we split logmFD into average and change in logmFD (p = `r round(p_mean_BMI,4)`) and there was no significant association of mean or change in logmFD with DMN FC.

@@

test_vif=lm(mean_Rew_conn ~ mean.BMI + within.BMI + mean.logmFD + within.logmFD +  Age_BL + Sex, data=final_FC)
vif_res=vif(test_vif)
``` 
As expected from the whole-brain findings, there was no association of mean BMI or within-subject BMI change and reward network connectivity in models adjusting for age, sex and logmFD.

<!-- [OLD stuff] GM mask: Group*Time Interaction -->
<!-- ### DMN for two timepoints
BMIONLYIG 
Upon inclusion of only a subsample entailing participants from the intervention group but not the waiting control group, there was still no effect for average BMI, but a significant effect of change in BMI on co-activation between the seed region NAcc and a voxel ($-18,-42,-9$ [mm])<!-- belonging to a cluster of $k = 2$ with $Z = 4.64$ ($p_{FWEcorr} = 0.049$) (in GSR corrected connectivity maps). However the effect did not survive recalculation with Wild bootstrap ($p_{FWEcorr} = 0.426$)
<!-- BMI2TP 
Running the tests on a subsample of only two time points and omitting data from the third did not yield any significant co-activation with the ROI for neither of the preprocessing variants. Omitting mean FD did not change these results. -->
<!-- GM mask 
Results obtained from analyses with the grey matter masks were equal to those reported for the brain masks except test statistics for the above reported significant longitudinal effect for the NAcc in the intervention group differed only marginally ($q_{FDRcorr} = 0.048$) and did not survive recalculation with Wild bootstrap either ($q_{FDRcorr} = 0.552$).
In regard to the exploratory expeditions on subsamples, consistent with previous results, no evidence for co-activation was found when using the grey matter mask.
 --> 
<!-- Parametric and non-parametric significance test for the model testing for time-by-group interaction did not show any different results when using a grey matter mask. Again, parametric estimation and non-parametric estimation differed in regard to the time-by-group interaction of PCC and voxel ($18,42,30$) in the connectivity maps without GSR correction. While co-activation was significant in parametric estimation ($Z = 4.72$, $q_{FDRcorr} = 0.040$), it was non-significant when using wild bootstrap ($q_{\mathrm{FDRcorr}} = 0.998$)<!-- also FWE-corr n.s -->

### Effect of FD (exploratory analysis)
#### Whole-brain analysis
To investigate the unique contributions of the average value of and the longitudinal change in mean FD, we constructed two additional models. 
First, BMI measures are excluded from the design matrix. Covariance of the mFD and outcome meausres is investigated by splitting mFD into average mFD and longitudinal mFD change and by inspecting associations with each predictor. Age and sex were nuisance covariates.
<!-- #### Reward -->
As shown in Table \@ref(tab:tableFDmodel), there is a significant co-activation of one cluster with the NAcc. Cluster co-activation appeared irrespective of the preprocessing procedure, differing mainly in cluster size.
<!-- ### DMN -->
We did not find a any clusters for functional connectivity with the PCC.

<!--### Effect on hypothalamic connectivity (exploratory analysis)-->

```{r tableBMImodel, echo=TRUE}
source("report/tables.R")
TableList <- mk_FCTables()
tab_BMImodel <- TableList[[1]]
tab_BMImodel
```

```{r tableBMIFDmodel, eval=FALSE}
tab_BMIFDmodel <- mk_FCTables()[[3]]
tab_BMIFDmodel
```

```{r tableFDmodel, eval=FALSE}
tab_FDmodel <- mk_FCTables()[[2]]
tab_FDmodel
```

>>>>

```{r "knit setup", include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE) # fig.height=2, fig.width=4)
options(knitr.table.format = "latex")
options(knitr.kable.NA = '') # hide NA in tables
# set working directory of knit (RMarkdown) to directory of Rproj
opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
setwd(knitr::opts_knit$get("root.dir"))

source("report/figures.R")
source("report/tables.R")
source("report/create_sample_df.R")
```

@@

Resting state functional magnetic resonance imaging (rsfMRI) is a powerful technique to investigate the dynamic organization of the brain via the correlation of spontaneous neural activity. Anatomically distinct brain regions show functional connectivity (FC), i.e. correlated neural activity over time, in the absence of a cognitive task. Crucially, these brain networks correspond to networks activated during specific cognitive processes, e.g. visual, auditory and somatosensory processing, or higher-order cognitive functions, i.e. attention and executive control [@Smith_2009], and can thus be seen as default configuration of the brain. In the context of bariatric surgery, the default mode network, a higher-order network, involved in interoception and governing shifts between external-internal processes, and the reward network, processing hedonic value and internal motivation, are promising candidates to mediate altered gut-brain signaling.

@@

Higher BMI and obesity have been associated with differences in DMN FC in several resting state and task-based fMRI studies [@Beyer_2017; @Kullmann_2011; @Chao_2018; @Wijngaarden_2015, @Borowitz_2020; @Ding_2020; @Doucet_2017; @Sadler_2018]. 

@@

```{r "Load data", include=FALSE}
final <- create_sample_df(group = "both", tp = "all")

@@

```{r figBMIdescr, fig.align='center', out.width="90%", fig.cap="BMI over time.", echo=FALSE}
# eval = TRUE, fig.asp = 0.7, fig.width = 6
fig_BMIdescr <- mk_figBMIdescr(final)
fig_BMIdescr
```

@@

<!-- MODEL SETUP --> In a first group of models, we tested the interaction of time and group $FC = group + time + group*time$ as specified in the preregistration (*We will use the SwE toolbox (implemented in SPM 12.7486 run in MATLAB version > 9.0) to calculate the group-by-timepoint interaction*)
We did not perform the analysis as preregistered (i.e. without covariates and additional sensitivity analyses), but adjusted for baseline age and sex (Model 1a), and baseline age, sex and the logarithm of mean FD (Model 1b). 
<!-- DESIGN MATRIX, vielleicht beide auch als Abbildung kurz zeigen? --> In this first model, time was represented as factor because we did not consider the effect of time to be strictly linear, and thus continuous modeling of time, inappropriate (see Figure \@ref(fig:figDesignmatrix)). The resulting factorial design <!-- check: flexible factorial? --> contained one regressor for each time point per group.
```{r figDesignmatrix, fig.align = 'center', fig.ncol = 2, out.width = "50%", fig.subcap=c('Design matrix with time as categorial factor.','Design matrix with time as continuous variable.'), fig.cap="Possible design matrices of time-by-group model.", echo=FALSE}
# fig.show="hold",  out.width="40%", fig.height = 8,
PlotList <- mk_figDesignMatrix()

@@

## Quality control
We performed different quality controls for the rsfMRI data.
First, we inspected parameters of head motion (mFD) and a measure of whole-brain signal variation which is reflective of noise in the data (DVARS).
As described in @Beyer_2020, there was a group-by-time interaction on mFD. The intervention group reduced in head motion in the follow-up assessments compared to the control group. 
Further, DVARS did not qualitatively differ between groups and time points, and the correlation of DVARS and mFD remained similar over time points.
We concluded that even though there was a decrease in head motion in the intervention versus control group, the impact of head motion on imaging data remained similar over the study period in both groups.
```{r tableQcfc, echo=FALSE}
res <- read.csv("report/FDFCcorrelations.csv")
rownames(res) <- c("minimally processed", "AROMA", "AROMA + CC", "AROMA + CC + GSR")
colnames(res) <- c('mean FD-QC', 'median FD-QC', 'sig. vertex', 'sig. vertex BH', 'distance-FD-QC', 'pval')

@@

```{r FDFC correlation, eval=FALSE, echo=FALSE}
# do not evaluate if saved already
source("/data/pt_02161/Analysis/Preprocessing/qa/rs_qa/group_level_QA/QC_FC_correlations/calc_qc_fc_correlations.R")
FDFC=calc_qc_fc(final)
write.csv(FDFC,"report/FDFCcorrelations.csv", row.names = FALSE)
```

```{r prepare tsnr ROI plot, eval=FALSE}
final$subj.ID_tp=paste0(final$subj.ID,'_', final$tp)

@@

p<-ggplot(aes(as.factor(roi), mean),data=final_tsnr) +
  geom_violin(aes(fill=roi), alpha=0.5) + geom_jitter(height = 0, width = 0.1) + xlab("Region of interest") + ylab("average tSNR") + 
  scale_fill_manual(values=wes_palette("Darjeeling2",5,type="discrete")[2:3]) +
  theme(axis.text=element_text(size=10),
  axis.title=element_text(size=12),
  strip.text = element_text(size=12),
  legend.position="")
ggsave("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/tsnr.jpeg", height = 6.604, units="cm", p)
```

```{r figNetworks, echo=FALSE, out.width="33%", out.height="30%", fig.cap="Resting state networks based on AROMA+CC+GSR input data, one-sample t-tests adjusting for age and sex over all timepoints and participants with bootstrapped clusterwise inference (p FWE<0.05)", fig.subcap=c("DMN seeded from precuneus", "Reward network seeded from Nucleus Accumbens", "tSNR differences between regions"),fig.show='hold',fig.align='center'}
#networks=c("./report/DMN_thresholded_for_clusterFWE005.png","./report/Rew_thresholded_for_clusterFWE005.png")
networks=c("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/DMN_thresholded_for_clusterFWE005.png", "/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/Rew_thresholded_for_clusterFWE005.png")
tsnr=c("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/tsnr.jpeg")
knitr::include_graphics(networks)
knitr::include_graphics(tsnr)
```

Figure \@ref(fig:figNetworks) shows the DMN and reward network average maps which we used for extracting aggregated FC values. Both networks included the typically described brain regions (i.e. medial prefrontal cortex, parietal lobules and posterior hippocampus for the DMN, and anterior cingulate, right amygdala and ventromedial-prefrontal cortex for the reward network). 
Yet, the reward network was less pronounced (overall lower T-values) and symmetric (no significant connectivity with left amygdala), which might be due to lower SNR in the seeded brain region (see Figure \@ref(fig:figNetworks)).

@@

<!-- Brain mask: Group*Time Interaction [CORRECTED] -->
Against our initial hypothesis, there was no interaction effect of group and time point for FC for neither the PCC nor the NAcc seeds. There also was no significant main effect for any of the covariates of interest (time, group) in clusterwise inference with FWE correction.

####  Model 1b (with adjustment for age and sex, meanFD)
Similarly, there were no significant clusters when including mFD. Results did not differ between AROMA+CC and AROMA+CC+GSR preprocessing pipelines.

@@

## Interaction effect group by time on aggregated FC (preregistered analysis)
```{r load agg FC data and prepare, echo=FALSE}
agg_FC_DMN=read.table("/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/aggFC/DMN/agg_FC_DMN.txt")
agg_FC_DMN_ID=read.table("/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/aggFC/DMN/agg_FC_DMN_ID.txt")
agg_FC_Rew_ID=read.table("/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/aggFC/Rew/agg_FC_Rew_ID.txt")
agg_FC_Rew=read.table("/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/aggFC/Rew/agg_FC_Rew.txt")
for (i in c(1:nrow(agg_FC_DMN_ID))){
  agg_FC_Rew[i,"subj.ID_tp"]=strsplit(base::strsplit(as.character(agg_FC_Rew_ID[i,"V1"]),'/')[[1]][10],'_nacc')[[1]][1]
  agg_FC_DMN[i,"subj.ID_tp"]=strsplit(base::strsplit(as.character(agg_FC_DMN_ID[i,"V1"]),'/')[[1]][10],'_pcc')[[1]][1]
}

@@

#calculate mean FD over timepoints
final_FC.meanFD=tapply(X=final_FC$logmFD,
                        INDEX=final_FC$subj.ID, FUN=mean, na.rm=TRUE)
final_FC$mean.logmFD=
  final_FC.meanFD[as.numeric(final_FC$subj.ID)]
final_FC$within.logmFD=
  final_FC$logmFD-final_FC$mean.logmFD
```
```{r figaggDMN, fig.cap="DMN connectivity per group over time", out.width="75%", echo=FALSE}
fig_DMNconn<-mk_figDMNdescr(final_FC)
fig_DMNconn
```
```{r figaggRew,fig.cap="Reward network connectivity per group over time", out.width="75%", echo=FALSE}
fig_Rewconn<-mk_figRewdescr(final_FC)
fig_Rewconn
```

```{r identify outliers in these plots, echo=FALSE}
### SUMMARY: all in all there is no reason to exclude the subjects with outlying values in mean Rew/DMN connectivity
final_FC[final_FC$mean_DMN_conn<0.1, c("mean_DMN_conn", "mean_Rew_conn", "subj.ID_tp", "subj.ID", "tp", "condition")] 
#ADI003_fu2 & ADI041_fu -> do not have particularly high head motion 
#plot(final_FC$mean_DMN_conn, final_FC$logmFD) -> see two left points
#carpet plots look ok, brain maps of DMN look okish, though not pronounced
#ADI003_fu2 & ADI041 fu: FS issues (Final_score:3, corrected)
qa_info=read.csv("/data/p_02161/ADI_studie/metadata/final_sample_MRI_QA_info.csv")
qa_info[qa_info$subj.ID=="ADI003"&qa_info$tp=="fu2",]
qa_info[qa_info$subj.ID=="ADI041"&qa_info$tp=="fu",]

final_FC[final_FC$mean_Rew_conn>0.3, c("mean_Rew_conn", "subj.ID_tp", "subj.ID", "tp", "condition")]
plot(final_FC$mean_Rew_conn, final_FC$logmFD)
#-> do not have particularly high head motion (two right points)
#ADI069_bl
#nifti looks ok, carpet plot shows lot of stripes.
qa_info[qa_info$subj.ID=="ADI069"&qa_info$tp=="bl",]
#FS issues, parts of skull included
#ADI111_fu2
#nifti looks ok, some residual carpet plot noise
qa_info[qa_info$subj.ID=="ADI111"&qa_info$tp=="fu2",]
#no FS issues
```

There was no significant time by group interaction effect for aggregated DMN and reward network FC (see Figures \@ref(fig:figaggDMN) and \@ref(fig:figaggRew) and supplements for a detailed summary of the models).

@@

@@

R1 = lme4::lmer(mean_Rew_conn ~ tp*group_factor + Age_BL + Sex + logmFD +(1|subj.ID), data=final_FC)
R0 = lme4::lmer(mean_Rew_conn ~ tp + group_factor + Age_BL + Sex +logmFD +(1|subj.ID), data=final_FC)
anvRew3tp=anova(R1,R0)
```
```{r Rew3tp results, caption="DMN connectivity compared bl to fu", results="asis", eval=FALSE}
#tab_model(R1,R0)
texreg(list(R0,R1), reorder.coef=c(1, 2, 3, 4, 8,9,5,6,7),custom.coef.names=c("Intercept (BL, NBARS)", "Timepoint FU", "Timepoint FU2", "Group BARS", "Age", "Sex","logmFD","Timepoint FU x Group", "Timepoint FU2 x Group"), custom.model.names = c("R0", "R1"), float.pos = "h",  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
```

## Exploratory Analyses
### Effect of mean and change in BMI 
#### Whole brain analysis
##### BMI-only model
Next, we tested for significant effects of the average BMI and the individual change in BMI longitudinally across the three time points.
<!-- ### DMN -->
<!-- cc -->
Using the images preprocessed with AROMA+CC, we found a significant negative association of average BMI and DMN connectivity in the lingual gyrus, mid orbital gyrus and temporal gyrus, which remained stable after including mFD as nuisance covariate (FWE cluster p<0.05). When using GSR preprocessed data, none of the clusters was significantly associated with average BMI.
```{r "BMIagesexPCCccavgBMIdeact",fig.cap="Negative association of average BMI and FC with PCC/precuneus in AROMA+CC preprocessed data", fig.subcap=c("adjusted for age and sex", "adjusted for age, sex and mFD"), fig.hold=TRUE, out.width = "50%"}
knitr::include_graphics(c("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/BMIagesex_PCCcc_avgBMI_deact.pdf", "/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/BMIagesexfd_PCCcc_avgBMI_deact.pdf"))
```

NAcc connectivity was not significantly related to neither mean nor change in BMI, for either of the preprocessing schemes. 
Results with significant cluster co-activation are depicted in Table \@ref(tab:tableBMImodel).

##### BMI and FD model
To disentangle effects of change in BMI and mFD in this dataset, we split mFD into average mFD and change in mFD. Age and sex were kept as nuisance covariates.
<!-- ### DMN -->
<!-- cc --> Average BMI remained significantly negatively associated with the three clusters above when splitting FD into its components. Results are detailed in Table \@ref(tab:tableBMIFDmodel).
Test statistics only differed slightly from the effect found when using mFD as nuisance covariate.
<!-- gsr -->As before, the effect did not survive when using GSR-corrected data.

```{r "BMIFDagesexPCCccavgBMIdeact", fig.cap="Negative association of average BMI and FC with PCC/precuneus in AROMA+CC preprocessed data, adjusted for BMI change, average mFD and change in mFD", fig.hold=TRUE, out.width = "100%"}
knitr::include_graphics("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/BMIFDagesex_PCCcc_avgBMI_deact.pdf")
```

<!-- ### Reward -->
<!-- cc -->
For the reward network, we found a significant negative association of change in BMI with FC between NAcc and a cluster in the posterior-medial frontal region, i.e. supplementary motor area (see Table \@ref(tab:tableBMIFDmodel)). Voxel activation at local maximum within this cluster was significant at peak level after FWE-correction (p=$0.028$), also when evaluating the GSR corrected data (p=$0.038$).
```{r "BMIFDagesexNACCcccgnBMIdeact", fig.cap="Significant cluster with a negative Effect of change BMI on NAcc to posterior-medial frontal region, adjusted for age, sex, mean BMI, average mFD and change in mFD", fig.subcap=c("in AROMA+CC", "in AROMA+CC+GSR data"), out.width = "50%", fig.hold=TRUE, eval=TRUE}
knitr::include_graphics(c("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/BMIFDagesex_NACCcc_cngBMI_deact.pdf","/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/BMIFDagesex_NACCgsr_cngBMI_deact.pdf"))
```
Moreover, FC between the NAcc and one cluster in a more posterior region, i.e. motor cortex, was significantly positively related to average mFD. This result remained similar when using GSR corrected data.
```{r "BMIFDagesexNACCccavgFDact", fig.cap="Significant cluster with a positive Effect of average mFD on connectivity between NAcc and motor cortex, adjusted for age, sex, mean BMI, change in BMI and change in mFD", fig.subcap=c("in AROMA+CC", "in AROMA+CC+GSR data"), out.width = "50%", eval=TRUE}
knitr::include_graphics(c("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/BMIFDagesex_NACCcc_avgFD_act.pdf","/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/BMIFDagesex_NACCgsr_avgFD_act.pdf"))
```

@@

p_mean_BMI=anvmeanBMI$`Pr(>Chisq)`
```
The association also remained significant when we split mFD into average and change in mFD (p = `r round(p_mean_BMI,4)`) and there was no significant association of mean or change in mFD with DMN FC.

@@

test_vif=lm(mean_Rew_conn ~ mean.BMI + within.BMI + mean.logmFD + within.logmFD +  Age_BL + Sex, data=final_FC)
vif_res=vif(test_vif)
``` 
As expected from the whole-brain findings, there was no association of mean BMI or within-subject BMI change and reward network connectivity in models adjusting for age, sex and mFD.

<!-- [OLD stuff] GM mask: Group*Time Interaction -->
<!-- ### DMN for two timepoints
BMIONLYIG
Upon inclusion of only a subsample entailing participants from the intervention group but not the waiting control group, there was still no effect for average BMI, but a significant effect of change in BMI on co-activation between the seed region NAcc and a voxel ($-18,-42,-9$ [mm])<!-- belonging to a cluster of $k = 2$ with $Z = 4.64$ ($p_{FWEcorr} = 0.049$) (in GSR corrected connectivity maps). However the effect did not survive recalculation with Wild bootstrap ($p_{FWEcorr} = 0.426$)
<!-- BMI2TP
Running the tests on a subsample of only two time points and omitting data from the third did not yield any significant co-activation with the ROI for neither of the preprocessing variants. Omitting mean FD did not change these results. -->
<!-- GM mask
Results obtained from analyses with the grey matter masks were equal to those reported for the brain masks except test statistics for the above reported significant longitudinal effect for the NAcc in the intervention group differed only marginally ($q_{FDRcorr} = 0.048$) and did not survive recalculation with Wild bootstrap either ($q_{FDRcorr} = 0.552$).
In regard to the exploratory expeditions on subsamples, consistent with previous results, no evidence for co-activation was found when using the grey matter mask.

<!-- Parametric and non-parametric significance test for the model testing for time-by-group interaction did not show any different results when using a grey matter mask. Again, parametric estimation and non-parametric estimation differed in regard to the time-by-group interaction of PCC and voxel ($18,42,30$) in the connectivity maps without GSR correction. While co-activation was significant in parametric estimation ($Z = 4.72$, $q_{FDRcorr} = 0.040$), it was non-significant when using wild bootstrap ($q_{\mathrm{FDRcorr}} = 0.998$)<!-- also FWE-corr n.s -->
